---
title: "Urothelium_Integrationof_UUOandIRI"
author: "Xin Wang"
date: "2024-02-22"
email: xin.wang@nationwidechildrens.org
output: html_document

Description:
  Method Part1: Select the Urothelium cells in the object
  Method reference: https://satijalab.org/seurat/articles/integration_introduction.html

Dataset Description: 

  uni-IRI and UUO on wild type adult male mice: which enthanized at multiple timepoints
  as the disease progress (0.6 house and 2, 7. 14 28 days post uni-IRI or 0,2,4, 6, 10,14 days post UUO), n=2 for each timepoint)
  GEO: GSE190887
  GEO reference: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi
  Reference: Li H, Dixon EE, Wu H, Humphreys BD. Comprehensive single-cell transcriptional profiling defines shared and unique epithelial injury responses during kidney fibrosis. Cell Metab 2022 Dec 6;34(12):1977-1998.e9.
  
  Platform: Illumina NovaSeq 6000
  
  snRNA-seq provied another strategy for performing single cell transcriptomics where individual nuclei instead of cells are captured and sequenced.
   Advantages: snRNA-seq over scRNA-seq is that they don't need dissociation to preseve cell integrity during sample preparation.
   Disadvantages: loss of transcripts are primarily located in the cytoplasm potentially limited the availiablity of biological signals for gene with little nuclear localization.

---

```{r}
module purge
ml GCC/9.3.0
ml OpenMPI/4.0.3
ml R/4.4.0
R
capabilities()
plot(1:20)
```

## Step 1: Load the packages
```{r}
## Please install all the packages before loading them.
library(Seurat)
library(gridExtra)
library(ggplot2)
library(tidyverse)
library(data.table)
#install.packages('R.utils')
library(R.utils)
library(reshape2)
library(stringr)
library(harmony)
library(cowplot)

library(patchwork)
library(decoupleR)

# Extra libraries
library(dplyr)
library(pheatmap)

library(patchwork)
# we use v3 instead of v5
#options(Seurat.object.assay.version = "v5")

options(future.globals.maxSize = 1e12)


# remotes::install_github("satijalab/seurat", "seurat5", quiet = TRUE)
# remotes::install_github("satijalab/seurat-data", "seurat5", quiet = TRUE)
# remotes::install_github("satijalab/azimuth", "seurat5", quiet = TRUE)
# remotes::install_github("satijalab/seurat-wrappers", "seurat5", quiet = TRUE)
# remotes::install_github("stuart-lab/signac", "seurat5", quiet = TRUE)
# #remotes::install_github("bnprks/BPCells", quiet = TRUE)
# remotes::install_github("bnprks/BPCells", quiet = TRUE)
# # Enter commands in R (or R studio, if installed)

sessionInfo()
```



## Step 2: Set up the workspace environment:
```{r}

setwd("/home/gdjacksonlab/lab/xxw004/UUO/Results/KUDOs/Integration/IntegratingUUO/")

UUODir<- c("/home/gdjacksonlab/lab/xxw004/UUO/Datasets/")
KUDODir <- c("/home/gdjacksonlab/lab/xxw004/UUO/Results/KUDOs/Integration/DoubletFinder/")
```


## we preprocessing UUO the datasets:
```{r}
## read the preprocessing
UUOProjectObject_F<- readRDS(paste0(UUODir,"UUP_Urothelium_ObjectMatrixTest.rds"))

MetaAnnotation<- read.csv(paste0(UUODir,"GSE190887_meta_cell_type_sample.csv"), header = T, row.names = 1)
rownames(MetaAnnotation)<-str_replace_all(rownames(MetaAnnotation), "\\.","_")
MetaAnnotation %>% head()
UUOProjectObject_F@meta.data
UUOProjectObject_F<-AddMetaData(object = UUOProjectObject_F, metadata = MetaAnnotation)
## we also add the meta data inside
UUOProjectObject_F@meta.data %>% head()

## add the mitchondiron percentage
#Test<- as.Seurat(UUOProject0,counts = "counts",project = "SingleCellExperiment",)
UUOProjectObject_F[['percent.mt']] <- PercentageFeatureSet(UUOProjectObject_F, pattern = "^mt[-\\.]")

## add the ribosome genes
UUOProjectObject_F[['percent.rRNA']] <- PercentageFeatureSet(UUOProjectObject_F, pattern = "^Rp[sl][[:digit:]]")

## add the condition 
nrow(UUOProjectObject_F@meta.data)

Idents(UUOProjectObject_F)

#DimPlot(UUOProjectObject_F)

#UUOProjectObject_F@meta.data %>% mutate(condition = stringr:: str_split(sample,"_", simplify =T)[,1], Day= stringr :: str_split(sample,"_",simplify =T)[,2])
UUOProjectObject_F@meta.data %>% mutate(condition = stringr:: str_split(sample,"_", simplify =T)[,1])

## extract the subset of Urothelium
levels(factor(UUOProjectObject_F@meta.data$celltype0421))
#Urothelium <-subset(UUOProjectObject_F, subset = celltype0421 == "Uro")
#Urothelium <-subset(UUOProjectObject_F, subset = Upk3a > 0 | Krt5 > 0 | Krt8 >0| Upk1b >0 |Krt14 >0 |Upk2 >0 |Sprr1a >0 |Psca >0 |Itga6>0 |Pparg >0)
table(UUOProjectObject_F@meta.data$sample)
```

# Extract the Urothlium
```{r}
# here we used two ways to confidently determine the Urothelium cells
Idents(UUOProjectObject_F) <- UUOProjectObject_F@meta.data$celltype0421

# We only check the expression of Upk genes and Krt genes
Idents(UUOProjectObject_F)

UUOProjectObject_F@meta.data <- UUOProjectObject_F@meta.data %>% mutate(Group= case_when(grepl("UUO",sample) ~"UUO", 
                                                 grepl("Health",sample) ~ "Health",
                                                grepl("IRI",sample) ~"IRI",
                                                .default = as.character(sample)
                                                 ))
# we only select the Urothelium that in Health and UUO model
UUOProjectObject_F_UUOHealth <- subset (UUOProjectObject_F, subset =(Group == "Health" | Group== "UUO") )

# setup identity
Idents(UUOProjectObject_F_UUOHealth) <- UUOProjectObject_F_UUOHealth@meta.data$sample

# reorder the samples 
Idents(UUOProjectObject_F_UUOHealth)<- ordered(factor(Idents(UUOProjectObject_F_UUOHealth)),levels=c("Health","UUO_Day2","UUO_Day4","UUO_Day6","UUO_Day10","UUO_Day14"))

nrow(table(UUOProjectObject_F_UUOHealth@meta.data$celltype0421))
# check the DotPlot

UrotheliumMarkerDotPlots<-DotPlot(UUOProjectObject_F_UUOHealth, features = c("Upk3a","Upk1a","Upk1b","Upk2", "Krt5","Krt14","Pparg"),col.min = -1.5,col.max = 1,cols="RdBu")  + guides(color = guide_colorbar(title = 'AverExp'),
         size = guide_legend(title = "PercExp")) + coord_flip() + RotatedAxis()  
UrotheliumMarkerDotPlots
ggsave("UrotheliumMarkerDotPlots.pdf",plot = UrotheliumMarkerDotPlots, height = 4, width = 5)
# plot the pparg
#DotPlot(UUOProjectObject_F_UUOHealth, features = "Pparg")
# FeaturePlot(UUOProjectObject_F_UUOHealth, features = "Pparg")

UUOInjuryMarkersDotPlot<- DotPlot(UUOProjectObject_F_UUOHealth, features = c("Havcr1","Lcn2","Col1a1","Col3a1","Acta2"),cols="RdBu")  + guides(color = guide_colorbar(title = 'AverExp'),
         size = guide_legend(title = "PercExp")) + coord_flip() + RotatedAxis()
ggsave("UUOInjuryMarkersDotPlot.pdf",plot = UUOInjuryMarkersDotPlot, height = 3.5, width = 4)
```

# We read the KUDO preproceeded data
```{r}

KudoObj<- readRDS(paste0(KUDODir,"Vehicle_withdoublets.rds"))

KudoFilObj <- subset(KudoObj, subset = doublet_id == "Singlet")

KudoObj@meta.data %>% head()
DimPlot(KudoFilObj, group.by = "seurat_clusters")
Idents(KudoFilObj) = "orig.ident"
# check the expression of kidney biomarkers in the KUDO dataset
markers_mouse_expanded_withothers<- c("Krt5","Krt14","Trp63",
  "Krt8","Krt18","Foxa1",
  "Upk1a","Upk1b","Upk2","Upk3b","Upk3a","Krt20",
  "Lrp2","Slc34a1","Slc13a3", # Proximal tubule cells (PT)
  "Slc12a3","Pvalb","Wnk1",# Distal connecting tublue (DCT)
"Umod","Slc12a1","Cldn10", # Thick ascending limb cells (TAL)
                      "Klk1","Slc8a1","Calb1", # Connecting tubule (CNT)
                     "Slc14a2","Fst","Bst1", # descending thin  limb cells (DTL)
                     #"Akr1b1","Sh3gl3","Prox1", # ascending thin limb cells (ATL)
                     #"Enox1","Thsd4", #Macula densa
                      "Aqp2","Hsd11b2","Scnn1g", # principal cells (PC)
                    "Atp6v1g3","Aqp6","Slc26a7", # Intercalated cells
                    "Cdh5","Igfbp3","Pecam1", # Endothelium
                    "Col3a1","Fbn1","Lum","Col1a1", # Fibroblast
                    "Ireb2","Alas2", # Plasmacytoid dentritic cells
                    "S100a8","S100a9","Il1b", # Neutrophil
                    "C1qa","C1qb","Aif1", # Macrophage
                    "Cxcr6","Cd247","Cd3e", # T cells
                    "Igkc","Cd79a","Cd79b", # B cells
                    "Ccl5","Nkg7","Cd7" # NK/Cd8 cells
  
  )
KidneyMarkersDotPlot<- DotPlot(KudoFilObj, features = markers_mouse_expanded_withothers) + RotatedAxis()
ggsave("KidneyMarkersDotPlot_Vehicle_MouseH_Reclustered.pdf", plot=KidneyMarkersDotPlot, height=4, width=20)

avg_exp <- AverageExpression(
  object = KudoFilObj,
  features = markers_mouse_expanded_withothers,
  #group.by = "celltype"   # or "seurat_clusters", "condition"
)
avg_exp%>% head()
# set up the order of the columns in the same order as the markers
aveg_exp_df <- as.data.frame(avg_exp$RNA)

# identify missing genes
missing_genes <- setdiff(markers_mouse_expanded_withothers, rownames(aveg_exp_df))
if (length(missing_genes) > 0) {
  zero_mat <- matrix(
    0,
    nrow = length(missing_genes),
    ncol = ncol(aveg_exp_df),
    dimnames = list(missing_genes, colnames(aveg_exp_df))
  )
  
  aveg_exp_df <- rbind(aveg_exp_df, zero_mat)
}

aveg_exp_df$Gene <- rownames(aveg_exp_df)

library(ggplot2)
# set the order of the genes in the same order as the markers
aveg_exp_df$Gene <- factor(aveg_exp_df$Gene, levels = markers_mouse_expanded_withothers)

AvereageExpressionBarPlot <-
ggplot(aveg_exp_df, aes(x = Gene, y = V1)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  theme_classic(base_size = 14) +
  labs(y = "Average Expression", x = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )
ggsave("AvereageExpressionBarPlot_Vehicle_MouseH_Reclustered.pdf", plot=AvereageExpressionBarPlot, height=4, width=12)


# we check the QC of the KUDO dataset
 QCReport <-
   VlnPlot(
     KudoFilObj,
     features = c("rDNA", "percent.mt", "nCount_RNA", "nFeature_RNA") ,
     pt.size=0,
     ncol = 1
   )
# save into pdf file
ggsave("QCReport_Vehicle_MouseH_Reclustered.pdf", plot=QCReport, height=8, width=4)
#getwd()
```


```{r}
# we compared the KUDO with Mouse Urothelium
# merge the Kudo Single cell and Mouse Urothelium
# delete the objects:
rm(KudoObj)
rm(UUOProjectObject_F_UUOHealth)

scrna <-
  merge(
    x = KudoFilObj,
    y = UUOProjectObject_F,
    project = "KudoUUOComparision"
  )

# add the meta data for the UUO and KUDO
scrna@meta.data %>% head()
table(scrna@meta.data$sample)
# change the sample name NA as the KUDOVehicle
scrna$sample[is.na(scrna$sample)] <- "KudoVehicle"
# change the condition NA as the KudoVehicle
scrna$Group[is.na(scrna$Group)] <- "KudoVehicle"

# check the number of cells in each sample(x, size, replace = FALSE, prob = NULL)
table((scrna$Group))
table((scrna$sample))
length(scrna$Group)
nrow (scrna@meta.data)

scrna <- scrna %>%  
  NormalizeData() %>% 
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
  ScaleData() %>% 
  SCTransform(vars.to.regress = c("percent.mt"))

## find most vaible features across sample to integrate


scrna <-  RunPCA(scrna, features = VariableFeatures(object = scrna),assay = "SCT",npcs=20)
saveRDS(scrna, file = "KudoUUOComparision_Scrna_PCA.rds")
#scrna <-readRDS(file = "KudoUUOComparision_Scrna_PCA.rds")
# then use the Harmony to remove batches
scrna <- RunHarmony(scrna, group.by.vars = 'sample', dims.use = 1:20, plot_convergence= FALSE, reduction = "pca", assay.use = "SCT", reduction.save = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_Harmony.rds")
scrna <- RunUMAP(scrna, dims = 1:20,reduction = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_RunUMAP.rds")
scrna <- FindNeighbors(scrna, dims = 1:20, reduction = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_FindNeighbor.rds")
scrna <- FindClusters(scrna, resolution = 0.2)

DimPlot(scrna, group.by = "sample")

saveRDS(scrna, file = "KudoUUOComparision_Scrna.rds")
FeaturePlot(scrna, dims = "umap",features = c("Havcr1","Lcn2","Col1a1"))
scrna<- readRDS("KudoUUOComparision_Scrna.rds")
```


```{r}
# we used the clustered method to extract the urothelium
Urothelium_S1<- subset(UUOProjectObject_F, idents = "Uro")
# 

Urothelium_S1@meta.data <- Urothelium_S1@meta.data %>% mutate(Group= case_when(grepl("UUO",sample) ~"UUO", 
                                                 grepl("Health",sample) ~ "Health",
                                                grepl("IRI",sample) ~"IRI",
                                                .default = as.character(sample)
                                                 ))
factor(Urothelium_S1@meta.data$Group)
Urothelium_S1@meta.data$Group<- ordered(factor(Urothelium_S1@meta.data$Group),levels=c("Health","IRI","UUO"))

# we only select the Urothelium that in Health and UUO model
Urothelium_S2 <- subset (Urothelium_S1, subset =(Group == "Health" | Group== "UUO") )

table(UUOProjectObject_F@meta.data$sample)
# we measure the Upk3a, Upk1a, Upk1b and Upk2 positive cell numbers
Upk3aUro <- subset(Urothelium_S2, subset = Upk3a > 2 )
Upk1aUro <- subset(Urothelium_S2, subset = Upk1a > 2 )
Upk1bUro <- subset(Urothelium_S2, subset = Upk1b > 2 )
Upk2Uro <- subset(Urothelium_S2, subset = Upk2 > 2 )
table(Upk3aUro@meta.data$sample)
table(Upk1aUro@meta.data$sample)
table(Upk1bUro@meta.data$sample)
table(Upk2Uro@meta.data$sample)

Krt5Uro <- subset(Urothelium_S2, subset = Krt5 > 2 )
Krt14Uro <- subset(Urothelium_S2, subset = Krt14 > 2 )

table(Krt5Uro@meta.data$sample)
table(Krt14Uro@meta.data$sample)
# we used the marker method to define the accurate urothelium

Urothelium <-subset(Urothelium_S2, subset = Upk3a > 2 | Upk1a >2 |Upk1b >2 |Upk2 > 2 )

table(Urothelium@meta.data$Group)
# Get the average expression of the genes
UrotheliumAverageExpression<- AverageExpression(object = Urothelium)$RNA
# Get the percentage of the genes
# get all the expressed genes 
UrotheliumExpressedGenes<- rownames(Urothelium)
#UrotheliuGenePercent <- PercentExpressing(seurat_object = Urothelium, features = UrotheliumExpressedGenes)

# other ways
plot <- DotPlot(object = Urothelium, features = UrotheliumExpressedGenes)
plot_data <- plot$data %>% 
  select(avg.exp,pct.exp, id)
# write them into file

write.csv(plot_data, file="Urothelium_ExpressedGenes_All_Ranking.csv")

table(Urothelium@meta.data$sample)
table(UUOProjectObject_F@meta.data$sample)

table(Urothelium_Upk3a_krt5$sample)
dim(Urothelium_Upk3a_krt5)
table(Urothelium@meta.data$sample)
dim(Urothelium)

```

```{r}
# We collected the dataset from the KUDO and Compared the expression of genes from Control
# Correlation of Average Expression
# 1. Compute average expression per cluster (organoid urothelium vs mouse urothelium).
# Pearson or Spearman correlation:
MouseCells <- subset(scrna, subset = Group == "Health")

# we only concentrate on the urothelium
MouseUrothelum@meta.data %>% head()
# 
MouseUrothelum <- subset(MouseCells, subset = celltype0421 == "Uro")


Idents(MouseUrothelum)
OrganoidUrothelium <- subset(scrna, subset = Group == "KudoVehicle")
# integrated the layers
DefaultAssay(OrganoidUrothelium) <- "RNA"

OrganoidUrothelium <- JoinLayers(OrganoidUrothelium)

# get the average expression of the genes in the mouse urothelium and organoid urothelium
expr_mat <- GetAssayData(
  MouseUrothelum,
  assay = "RNA",
  layer = "data"
)
mouse_avg <- Matrix::rowMeans(expr_mat)


Oexpr_mat <- GetAssayData(
  OrganoidUrothelium,
  assay = "RNA",
  layer = "data"
)
organoid_avg <- Matrix::rowMeans(Oexpr_mat)

# # Define a threshold for expressed genes
# expression_threshold <- 0  # Adjust this threshold as needed

# # Filter for expressed genes in both datasets

# expressed_genes_mouse <- names(mouse_avg[mouse_avg > expression_threshold])
# expressed_genes_organoid <- names(organoid_avg[organoid_avg > expression_threshold])

# Find shared expressed genes
shared_genes <- intersect(names(mouse_avg), names(organoid_avg))

length(shared_genes)
# Subset the datasets to only include shared expressed genes
mouse_avg_2 <- mouse_avg[shared_genes]
organoid_avg_2 <- organoid_avg[shared_genes]


# calculate the correlation between two datasets
cor_val <- cor(organoid_avg_2, mouse_avg_2, method="spearman")

library(reshape2)
library(ggplot2)

# Combine data into a data frame
data <- data.frame(mouse_avg_2, organoid_avg_2)

# Create the correlation dot plot
MouseUro_OrganoidUro_Cor_Plot<- ggplot(data, aes(x = log10(mouse_avg_2), y = log10(organoid_avg_2))) +
  geom_point(color = "gray", size = 3) +
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  labs(
    title = "Correlation Dot Plot",
    x = "Mouse Avg 2 (log2)",
    y = "Organoid Avg 2 (log2)"
  ) +
  theme_minimal() 
ggsave("MouseUro_OrganoidUro_Cor_Plot.pdf",plot = MouseUro_OrganoidUro_Cor_Plot, height = 4, width = 5)

# we subset the subset(scrna, subset = Group == "Health")

Urothelium_KUDO_MouseH_Obj <- subset(scrna, subset =  Group == "KudoVehicle" |(Group == "Health" & celltype0421 == "Uro"))
Urothelium_KUDO_MouseH_Obj@meta.data %>% head() 
DefaultAssay(Urothelium_KUDO_MouseH_Obj) <- "RNA"
Urothelium_KUDO_MouseH_Obj <- JoinLayers(Urothelium_KUDO_MouseH_Obj)

saveRDS(Urothelium_KUDO_MouseH_Obj, file = "Urothelium_KUDO_MouseH_Obj.rds")



  # +
  # annotate(
  #   "text",
  #   x = min(log10(mouse_avg_2)) + 0.5,
  #   y = max(log10(organoid_avg_2)) - 0.5,
  #   label = paste("Spearman Correlation:", round(cor_val, 2)),
  #   color = "black",
  #   size = 4,
  #   hjust = 0
  # )

# 2. Marker gene expression Overlaping
DotPlot(scrna, features = c("Krt5","Tp63","Upk1a","Upk2","Upk3a")) + RotatedAxis()
# Marker gene overlap (Jaccard index).
# J=∣Markersmouse​∪Markersorganoid​∣∣Markersmouse​∩Markersorganoid​∣​
Idents(Urothelium_KUDO_MouseH_Obj) <- "sample"
levels(Idents(Urothelium_KUDO_MouseH_Obj))
DotPlot(Urothelium_KUDO_MouseH_Obj, features = c("Krt5","Trp63","Upk1a","Upk2","Upk3a")) + RotatedAxis()
```

# method 2: the lineage marker expression correlation
```{r}
# Checking lineage marker expression correlation will tell you whether the urothelial identity program is preserved, independent of stress/proliferation noise.
markers_mouse <- c(
  "Krt5","Krt14","Trp63","Itga6","Krt17",
  "Krt8","Krt18","Foxa1","Gata3","Pparg",
  "Upk1a","Upk1b","Upk2","Upk3b","Upk3a","Krt20"
)
markers_mouse_expanded <- c(
  # Basal
  "Krt5","Krt14","Trp63","Itga6","Krt17","Col17a1","Sox2","Ngfr","Cd44","Lgr5","Pdpn","Epcam",
  # Intermediate
  "Krt8","Krt18","Foxa1","Gata3","Pparg","Grhl3","Upk1a","Upk1b","Upk2","Upk3a","S100a1","Cdh1","Tjp1",
  # Umbrella
  "Krt20","Shh","Plk2","Umod","Lypd3","Cldn8","Cldn4","Ocln"
)
# Subset the datasets to only include shared expressed genes
mouse_avg_markers <- mouse_avg[markers_mouse]
organoid_avg_markers <- organoid_avg[markers_mouse]
cor(mouse_avg_markers,
    organoid_avg_markers,
    method = "pearson")

cor(mouse_avg_markers,
    organoid_avg_markers,
    method = "spearman")

```

```{r}
# we selected the top 1000 variable genes in the mouse urothelium and organoid urothelium
mouse_top1000 <- head(VariableFeatures(MouseUrothelum), 1000)
organoid_top1000 <- head(VariableFeatures(OrganoidUrothelium), 1000)
# calculate the Jaccard index
intersection <- length(intersect(mouse_top1000, organoid_top1000))
union <- length(union(mouse_top1000, organoid_top1000))
jaccard_index <- intersection / union

```


```{r}
# we create a method for the measure the two objects average expresssion correlation
library(Seurat)
library(Matrix)
library(ggplot2)

# Function to compute correlation and generate dot plot
correlation_dotplot <- function(obj1, obj2, assay="RNA", slot="data",
                                genes=NULL, method="spearman",
                                log_scale=TRUE, plot_title=NULL,
                                save_plot=NULL) {
  
  # Get average expression
  expr1 <- Matrix::rowMeans(GetAssayData(obj1, assay=assay, slot=slot))
  expr2 <- Matrix::rowMeans(GetAssayData(obj2, assay=assay, slot=slot))
  
  # If a gene list is provided, subset
  if(!is.null(genes)) {
    shared_genes <- intersect(names(expr1), names(expr2))
    shared_genes <- intersect(shared_genes, genes)
  } else {
    shared_genes <- intersect(names(expr1), names(expr2))
  }
  
  # Subset to shared genes
  expr1_sub <- expr1[shared_genes]
  expr2_sub <- expr2[shared_genes]
  
  # Calculate correlation
  cor_val <- cor(expr1_sub, expr2_sub, method=method)
  
  # Prepare data frame for plotting
  plot_data <- data.frame(expr1_sub, expr2_sub)
  colnames(plot_data) <- c("Dataset1", "Dataset2")
  
  # Apply log10 if requested
  if(log_scale) {
    plot_data$Dataset1 <- log10(plot_data$Dataset1)
    plot_data$Dataset2 <- log10(plot_data$Dataset2)
    xlab <- paste0(deparse(substitute(obj1)), " Avg (log10)")
    ylab <- paste0(deparse(substitute(obj2)), " Avg (log10)")
  } else {
    xlab <- paste0(deparse(substitute(obj1)), " Avg")
    ylab <- paste0(deparse(substitute(obj2)), " Avg")
  }
  
  # Generate plot
  p <- ggplot(plot_data, aes(x=Dataset1, y=Dataset2)) +
    geom_point(color="gray", size=3) +
    geom_smooth(method="lm", color="black", se=FALSE) +
    labs(title = plot_title %||% "Correlation Dot Plot", x=xlab, y=ylab) +
    theme_minimal()
  
  # Save plot if filename provided
  if(!is.null(save_plot)) {
    ggsave(save_plot, plot=p, height=4, width=5)
  }
  
  # Return results
  return(list(correlation=cor_val, plot=p, shared_genes=shared_genes))
}

# Example usage:
res <- correlation_dotplot(MouseUrothelum, OrganoidUrothelium,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_Cor_Plot.pdf")

# Print correlation value
res$correlation

# Display plot
res$plot
length(res$shared_genes)
```

# Define the function II: top 1000 variable genes for jaccard_index 
```{r}
library(Seurat)
library(VennDiagram)

# Function to compute Jaccard index and draw Venn diagram
topVariableGenes_Jaccard <- function(obj1, obj2, assay="RNA", slot="data",
                                     n=1000, plot_title=NULL, 
                                     save_plot=NULL, verbose=TRUE) {
  
  # 1️⃣ Find top variable genes
  hvgs1 <- head(VariableFeatures(FindVariableFeatures(obj1, assay=assay, selection.method="vst", nfeatures=n)), n)
  hvgs2 <- head(VariableFeatures(FindVariableFeatures(obj2, assay=assay, selection.method="vst", nfeatures=n)), n)
  
  if(verbose) {
    message("Number of HVGs selected: ", length(hvgs1), " and ", length(hvgs2))
  }
  
  # 2️⃣ Compute Jaccard index
  shared <- intersect(hvgs1, hvgs2)
  union_genes <- union(hvgs1, hvgs2)
  jaccard_index <- length(shared) / length(union_genes)
  
  if(verbose) {
    message("Shared genes: ", length(shared))
    message("Union genes: ", length(union_genes))
    message("Jaccard index: ", round(jaccard_index, 3))
  }
  
  # 3️⃣ Generate Venn diagram
  venn_plot <- venn.diagram(
    x = list(Object1 = hvgs1, Object2 = hvgs2),
    filename = NULL,
    fill = c("skyblue", "pink"),
    alpha = 0.5,
    cex = 1.5,
    cat.cex = 1.2,
    main = plot_title %||% paste0("Top ", n, " HVGs Venn Diagram")
  )
  
  # Save plot if needed
  if(!is.null(save_plot)) {
    png(save_plot, width = 600, height = 600)
    grid.draw(venn_plot)
    dev.off()
  }
  
  # Return results
  return(list(hvgs_object1=hvgs1,
              hvgs_object2=hvgs2,
              shared_genes=shared,
              jaccard_index=jaccard_index,
              venn_plot=venn_plot))
}

# Example usage
Urotheliumres <- topVariableGenes_Jaccard(MouseUrothelum, OrganoidUrothelium,
                                n=2000,
                                plot_title="Mouse vs Organoid Top HVGs",
                                save_plot="Mouse_Organoid_HVGs_Venn.png")

# View Jaccard index
Urotheliumres$jaccard_index
Urotheliumres$venn_plot
```

# #####################
# UPK lineage
# #####################
```{r}
# for the objects, we only selected the cells that showe expression of Upk3a, Upk1a, Upk1b and Upk2.
UPKUrotheliumMouse<- subset(MouseUrothelum, Upk3a > 1 | Upk1a > 1 | Upk1b > 1 | Upk2 > 1)
# we also selected the cells that showe expression of Upk3a, Upk1a, Upk1b and Upk2 in the organoid urothelium
UPKUrotheliumOrganoid<- subset(OrganoidUrothelium, Upk3a > 1 | Upk1a > 1 | Upk1b > 1 | Upk2 > 1)

# print the correlation:
UPKres <- correlation_dotplot(UPKUrotheliumMouse, UPKUrotheliumOrganoid,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid UPK lineage Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_UPKPos_Cor_Plot.pdf")

# Print correlation value
UPKres$correlation

# Display plot
UPKres$plot

length(UPKres$shared_genes)

# # Venn diagram and Jaccard index for the top variable genes in the UPK positive cells
UPKtopvariable <- topVariableGenes_Jaccard(UPKUrotheliumMouse, UPKUrotheliumOrganoid,
                                n=2000,
                                plot_title="Mouse vs Organoid Top UPKs",
                                save_plot="Mouse_Organoid_top2000_UPKPos_Venn.png")
                                
UPKtopvariable$venn_plot 
UPKtopvariable$jaccard_index                             
```

# #####################
# KRT5, Krt14 lineage
# #####################
```{r}
# for the objects, we only selected the cells that showe expression of Krt5 and Krt14.
KRT5UrotheliumMouse<- subset(MouseUrothelum, Krt5 > 1 |Krt14 > 1)
# we also selected the cells that showe expression of Krt5 and Krt14.
KRT5UrotheliumOrganoid<- subset(OrganoidUrothelium, Krt5 > 1 |Krt14 > 1))

# print the correlation:
KRTres <- correlation_dotplot(KRT5UrotheliumMouse, KRT5UrotheliumOrganoid,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid KRT5 lineage Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_KRT5Pos_Cor_Plot.pdf")

# Print correlation value
KRTres$correlation
KRTres$plot
length(KRTres$shared_genes)
# Venn diagram and Jaccard index for the top variable genes in the KRT5 and KRT14 positive cells
KRTtopvariable <- topVariableGenes_Jaccard(KRT5UrotheliumMouse, KRT5UrotheliumOrganoid,
                                n=2000,
                                plot_title="Mouse vs Organoid Top KRT",
                                save_plot="Mouse_Organoid_top2000_KRTPos_Venn.png")
KRTtopvariable$jaccard_index   
```

# ##############
# UPK,KRT lineage
# ##############
```{r}
# we selected the KRT and UPK positive cells in the mouse urotheloium

UPKKRTUrotheliumMouse<- subset(MouseUrothelum, subset = (Upk3a > 1 | Upk1a > 1 | Upk1b > 1| Upk2 > 1 | Krt5 > 1 |Krt14 > 1)) 

# we also selected the KRT and UPK positive cells in the organoid urotheloium
UPKKRTUrotheliumOrganoid<- subset(OrganoidUrothelium,(Upk3a > 1 | Upk1a > 1 | Upk1b > 1| Upk2 > 1 | (Krt5 > 1 |Krt14 > 1)) 
# print the correlation:
UPKKRTres <- correlation_dotplot(UPKKRTUrotheliumMouse, UPKKRTUrotheliumOrganoid,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid UPK+ or KRT+ Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_UPKKRT5Pos_Cor_Plot.pdf")

# Print correlation value
UPKKRTres$correlation
UPKKRTres$plot
length(UPKKRTres$shared_genes)

# Venn diagram and Jaccard index for the top variable genes in the UPK+ or KRT+ cells
UPKKRTtopvariable <- topVariableGenes_Jaccard(UPKKRTUrotheliumMouse, UPKKRTUrotheliumOrganoid,
                                n=2000,
                                plot_title="Mouse vs Organoid Top UPK KRT",
                                save_plot="Mouse_Organoid_top2000_UPKKRTPos_Venn.png")
```


```{r}
# check the annotation of scrna
Idents(scrna)
scrna@meta.data %>% head()
scrna@meta.data$celltype0421
```

# ############### ##############
# ############### ##############
# Reclustering the Urothelium_KUDO_MouseH_Obj to check the overlapping of the two datasets
# ############### ##############
# ############### ##############

```{r}
Urothelium_KUDO_MouseH_Obj<- readRDS("Urothelium_KUDO_MouseH_Obj.rds")
# we recluster the Urothelium_KUDO_MouseH_Obj
Urothelium_KUDO_MouseH_Obj <- Urothelium_KUDO_MouseH_Obj %>%  
  NormalizeData() %>% 
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
  ScaleData() %>% 
  SCTransform(vars.to.regress = c("percent.mt"))

## find most vaible features across sample to integrate


Urothelium_KUDO_MouseH_Obj <-  RunPCA(scrna, features = VariableFeatures(object = Urothelium_KUDO_MouseH_Obj),assay = "SCT",npcs=20)
#scrna <-readRDS(file = "KudoUUOComparision_Scrna_PCA.rds")
# then use the Harmony to remove batches
Urothelium_KUDO_MouseH_Obj <- RunHarmony(Urothelium_KUDO_MouseH_Obj, group.by.vars = 'sample', dims.use = 1:20, plot_convergence= FALSE, reduction = "pca", assay.use = "SCT", reduction.save = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_Harmony.rds")
Urothelium_KUDO_MouseH_Obj <- RunUMAP(Urothelium_KUDO_MouseH_Obj, dims = 1:20,reduction = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_RunUMAP.rds")
Urothelium_KUDO_MouseH_Obj <- FindNeighbors(Urothelium_KUDO_MouseH_Obj, dims = 1:20, reduction = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_FindNeighbor.rds")
Urothelium_KUDO_MouseH_Obj <- FindClusters(Urothelium_KUDO_MouseH_Obj, resolution = 0.2)

OverlappingDimPlot<- DimPlot(Urothelium_KUDO_MouseH_Obj)
ggsave("OverlappingDimPlot.pdf",plot = OverlappingDimPlot, height = 4, width = 5)
OverlappingDimPlotSplit<- DimPlot(Urothelium_KUDO_MouseH_Obj, split.by = "sample")
ggsave("OverlappingDimPlotSplit.pdf",plot = OverlappingDimPlotSplit, height = 4, width = 9)

saveRDS(Urothelium_KUDO_MouseH_Obj, file = "Urothelium_KUDO_MouseH_Obj_Reclustered.rds")

Urothelium_KUDO_MouseH_Obj <- readRDS("Urothelium_KUDO_MouseH_Obj_Reclustered.rds")
markers_mouse_expanded <- c(
  # Basal
  "Krt5","Krt14","Trp63","Itga6","Krt17","Col17a1","Sox2","Ngfr","Cd44","Lgr5","Pdpn","Epcam",
  # Intermediate
  "Krt8","Krt18","Foxa1","Gata3","Pparg","Grhl3","Upk1a","Upk1b","Upk2","Upk3a","S100a1","Cdh1","Tjp1",
  # Umbrella
  "Krt20","Shh","Plk2","Umod","Lypd3","Cldn8","Cldn4","Ocln"
)

Idents(Urothelium_KUDO_MouseH_Obj) <- "sample"
UrothelialMarkersDotPlot<- DotPlot(Urothelium_KUDO_MouseH_Obj, features = markers_mouse_expanded) + RotatedAxis()
ggsave("UrothelialMarkersDotPlot_Urothelim_KUDO_MouseH_Reclustered.pdf", plot=UrothelialMarkersDotPlot, height=3, width=12)
markers_mouse_expanded_withothers<- c("Krt5","Krt14","Trp63",
  "Krt8","Krt18","Foxa1",
  "Upk1a","Upk1b","Upk2","Upk3b","Upk3a","Krt20",
  "Lrp2","Slc34a1","Slc13a3", # Proximal tubule cells (PT)
  "Slc12a3","Pvalb","Wnk1",# Distal connecting tublue (DCT)
"Umod","Slc12a1","Cldn10", # Thick ascending limb cells (TAL)
                      "Klk1","Slc8a1","Calb1", # Connecting tubule (CNT)
                     "Slc14a2","Fst","Bst1", # descending thin  limb cells (DTL)
                     #"Akr1b1","Sh3gl3","Prox1", # ascending thin limb cells (ATL)
                     #"Enox1","Thsd4", #Macula densa
                      "Aqp2","Hsd11b2","Scnn1g", # principal cells (PC)
                    "Atp6v1g3","Aqp6","Slc26a7", # Intercalated cells
                    "Cdh5","Igfbp3","Pecam1", # Endothelium
                    "Col3a1","Fbn1","Lum","Col1a1", # Fibroblast
                    "Ireb2","Alas2", # Plasmacytoid dentritic cells
                    "S100a8","S100a9","Il1b", # Neutrophil
                    "C1qa","C1qb","Aif1", # Macrophage
                    "Cxcr6","Cd247","Cd3e", # T cells
                    "Igkc","Cd79a","Cd79b", # B cells
                    "Ccl5","Nkg7","Cd7" # NK/Cd8 cells
  
  )
KidneyMarkersDotPlot<- DotPlot(Urothelium_KUDO_MouseH_Obj, features = markers_mouse_expanded_withothers) + RotatedAxis()
ggsave("KidneyMarkersDotPlot_Urothelim_KUDO_MouseH_Reclustered.pdf", plot=KidneyMarkersDotPlot, height=4, width=20)

# check the average expression of the markers in the mouse urothelium and organoid urothelium

avg_exp <- AverageExpression(
  object = Urothelium_KUDO_MouseH_Obj,
  features = markers_mouse_expanded_withothers,
  group.by = "sample"   # or "seurat_clusters", "condition"
)
avg_exp%>% head()
# set up the order of the columns in the same order as the markers
aveg_exp_df <- as.data.frame(avg_exp$RNA)


# identify missing genes
missing_genes <- setdiff(markers_mouse_expanded_withothers, rownames(aveg_exp_df))
if (length(missing_genes) > 0) {
  zero_mat <- matrix(
    0,
    nrow = length(missing_genes),
    ncol = ncol(aveg_exp_df),
    dimnames = list(missing_genes, colnames(aveg_exp_df))
  )
  
  aveg_exp_df <- rbind(aveg_exp_df, zero_mat)
}

aveg_exp_df$Gene <- rownames(aveg_exp_df)

# change to the long type
aveg_exp_df_long <- aveg_exp_df %>%
  pivot_longer(cols = -Gene, names_to = "Sample", values_to = "AverageExpression")

library(ggplot2)
# set the order of the genes in the same order as the markers
aveg_exp_df_long$Gene <- factor(aveg_exp_df_long$Gene, levels = markers_mouse_expanded_withothers)

AvereageExpressionBarPlot <-
ggplot(aveg_exp_df_long, aes(x = Gene, y = AverageExpression, fill= Sample)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  theme_classic(base_size = 14) +
  labs(y = "Average Expression", x = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )
ggsave("AvereageExpressionBarPlot_Urothelim_KUDO_MouseH_Reclustered.pdf", plot=AvereageExpressionBarPlot, height=4, width=15)



```