---
title: "Urothelium_Integrationof_UUOandIRI_BladderOrganoid_Similarity"
author: "Xin Wang"
date: "2026-04-22"
email: xin.wang@nationwidechildrens.org
output: html_document

Description:
  Method Part1: Select the Urothelium cells in the object
  Method reference: https://satijalab.org/seurat/articles/integration_introduction.html
    -- Global Correlation of Average Expression for the shared genes between the three datasets (organoid urothelium, mouse urothelium, and bladder urothelium organoid).
    -- Marker gene expression Overlaping for the shared genes between the three datasets (organoid urothelium, mouse urothelium, and bladder urothelium organoid).
    -- Jaccard index for the top 1000 variable genes in the three datasets (organoid urothelium, mouse urothelium, and bladder urothelium organoid).
    -- Correlation of the average expression of the genes in the three datasets (organoid urothelium, mouse urothelium, and bladder urothelium organoid).
    -- Correlation of the average expression of the lineage markers in the three datasets (organoid urothelium, mouse urothelium, and bladder urothelium organoid).
    -- Annotation of different urothelium subtypes: basal, intermediate and umbrella cells in the three datasets (organoid urothelium, mouse urothelium, and bladder urothelium organoid).
    -- Functional enrichment analysis of the top 1000 variable genes in the three datasets (organoid urothelium, mouse urothelium, and bladder urothelium organoid).
    

Dataset Description: 

   uni-IRI and UUO on wild type adult male mice: which enthanized at multiple timepoints
   as the disease progress (0.6 house and 2, 7. 14 28 days post uni-IRI or 0,2,4, 6, 10,14 days post UUO), n=2 for each timepoint)
   GEO: GSE190887
   GEO reference: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi
   Reference: Li H, Dixon EE, Wu H, Humphreys BD. Comprehensive single-cell transcriptional profiling defines shared and unique epithelial injury responses during kidney fibrosis. Cell Metab 2022 Dec 6;34(12):1977-1998.e9.
  
   Platform: Illumina NovaSeq 6000
  
  snRNA-seq provied another strategy for performing single cell transcriptomics where individual nuclei instead of cells are captured and sequenced.
    Advantages: snRNA-seq over scRNA-seq is that they don't need dissociation to preseve cell integrity during sample preparation.
    Disadvantages: loss of transcripts are primarily located in the cytoplasm potentially limited the availiablity of biological signals for gene with little nuclear localization.

    Urothelium organoid dataset: GSE131909_RAW
  Reference: Santos, C.P., Lapi, E., Martínez de Villarreal, J. et al. Urothelial organoids originating from Cd49fhigh mouse stem cells display Notch-dependent differentiation capacity. Nat Commun 10, 4407 (2019). https://doi.org/10.1038/s41467-019-12307-1n.

---

```{r}
module purge
ml GCC/9.3.0
ml OpenMPI/4.0.3
ml R/4.4.0
R
capabilities()
plot(1:20)
```

## Step 1: Load the packages
```{r}
## Please install all the packages before loading them.
library(Seurat)
library(gridExtra)
library(ggplot2)
library(tidyverse)
library(data.table)
#install.packages('R.utils')
library(R.utils)
library(reshape2)
library(stringr)
library(harmony)
library(cowplot)

library(patchwork)
library(decoupleR)

# Extra libraries
library(dplyr)
library(pheatmap)

library(patchwork)

#options(Seurat.object.assay.version = "v5")

options(future.globals.maxSize = 1e12)


# remotes::install_github("satijalab/seurat", "seurat5", quiet = TRUE)
# remotes::install_github("satijalab/seurat-data", "seurat5", quiet = TRUE)
# remotes::install_github("satijalab/azimuth", "seurat5", quiet = TRUE)
# remotes::install_github("satijalab/seurat-wrappers", "seurat5", quiet = TRUE)
# remotes::install_github("stuart-lab/signac", "seurat5", quiet = TRUE)
# #remotes::install_github("bnprks/BPCells", quiet = TRUE)
# remotes::install_github("bnprks/BPCells", quiet = TRUE)
# # Enter commands in R (or R studio, if installed)

sessionInfo()
```



## Step 2: Set up the workspace environment:
```{r}

setwd("/home/gdbecknelllab/xxw004/gdjacksonlab/UUO/Results/KUDOs/Integration/IntegratingUUO_Organoid/")
KUDODir <- c("/home/gdbecknelllab/xxw004/gdjacksonlab/UUO/Results/KUDOs/Integration/IntegratingUUO_Organoid/")

# Make directories for the results
# annotation directory
dir.create(paste0(KUDODir,"Annotation/"))
# DEG directory
dir.create(paste0(KUDODir,"DEG/"))
# Feature directory
dir.create(paste0(KUDODir,"Feature/"))
# Cell cell communication directory
dir.create(paste0(KUDODir,"CellCommunication/"))

```

```{r}

# Load the integrated object

# HarmonyIntegration
AllUrothelium<- readRDS("KudoUUOComparision_Scrna_Merged_RunHarmony.rds")

#AllUrothelium <- JoinLayers(AllUrothelium)

# rpca integration

```


```{r}
# ─── Pairwise correlation matrix for all 12 samples ───────────────────────────

# Step 1: Split AllUrothelium by Sample into a named list of Seurat objects
sample_levels <- c("Vehicle", "Yoda", "LOF", "GOF", "Health",
                   "UUO_Day2", "UUO_Day4", "UUO_Day6", "UUO_Day10", "UUO_Day14",
                   "GSM3827175_NMU_O_P", "GSM3827176_NMU_O_D")

DefaultAssay(AllUrothelium) <- "RNA"
AllUrothelium <- JoinLayers(AllUrothelium)

sample_list <- lapply(sample_levels, function(s) {
  subset(AllUrothelium, subset = Sample == s)
})
names(sample_list) <- sample_levels

# Step 2: Compute per-sample average expression vectors
avg_expr_list <- lapply(sample_list, function(obj) {
  Matrix::rowMeans(GetAssayData(obj, assay = "RNA", layer = "data"))
})

# Step 3: Find genes expressed (mean > 0) in ALL 12 samples
shared_genes <- Reduce(intersect, lapply(avg_expr_list, function(avg) {
  names(avg[avg > 0])
}))
message("Shared expressed genes across all 12 samples: ", length(shared_genes))

# Step 4: Build a gene × sample matrix restricted to shared genes
expr_matrix <- do.call(cbind, lapply(avg_expr_list, function(avg) avg[shared_genes]))
colnames(expr_matrix) <- sample_levels

# Step 5: Compute pairwise Spearman correlation matrix
cor_matrix <- cor(expr_matrix, method = "spearman")

# Step 6: Save the numeric matrix
write.csv(cor_matrix, "AllUrothelium_12Sample_SpearmanCorrelation.csv")

# Step 7: Heatmap
library(pheatmap)
pdf("AllUrothelium_12Sample_SpearmanCorrelation_Heatmap.pdf", width = 8, height = 7)
pheatmap(cor_matrix,
         color        = colorRampPalette(c("white", "#2166ac"))(100),
         breaks       = seq(0.8, 1, length.out = 101),
         display_numbers = TRUE,
         number_format   = "%.3f",
         fontsize_number = 7,
         cluster_rows = TRUE, cluster_cols = TRUE,
         main   = "Spearman Correlation — 12 Samples (shared expressed genes)",
         filename = "AllUrothelium_12Sample_SpearmanCorrelation_Heatmap.pdf",
         width = 8, height = 7)
dev.off()
```

# We tried to identify the best strategy for the integration of the datasets, we found that the Harmony integration can better integrate the datasets and show better overlapping of the samples in the UMAP plot. So we used the Harmony integration for the following analysis.

```{r}

# We read the director and identify the .rds files in the directory then run the loops to read the .rds files 
Filenames <- list.files(path = KUDODir, pattern = "KudoUUOComparision_Scrna_Merged_*.rds", full.names = TRUE, recursive = F)

# we read the .rds files and assign them to the objects, generate the files for the following analysis
for (i in 1:length(Filenames)) {
  
    # read the .rds file
    obj <- readRDS(Filenames[i])
    # assign the name of the object as the name of the file without the .rds extension
    name <- gsub(".rds", "", Filenames[i])
    # print the name of the object to confirm it was created
    print(paste("Loaded object:", gsub(".rds", "", basename(Filenames[i]))))
    # Join the layers
    obj <- JoinLayers(obj)
    # set up the object with Sample levels
    obj@meta.data$Sample <- ordered(factor(obj@meta.data$Sample), levels = c("Vehicle", "Yoda", "LOF", "GOF", "Health","UUO_Day2","UUO_Day4","UUO_Day6","UUO_Day10","UUO_Day14", "GSM3827175_NMU_O_P", "GSM3827176_NMU_O_D"))
    # generate the dimplot for the object with group.by Sample and split.by Sample
    
    AllUrotheliumDimPlotGroupBySample<- DimPlot(obj, group.by = "Sample")
    # save the plot 
    ggsave(paste0(name,"_DimPlotGroupBySample.pdf"),plot = AllUrotheliumDimPlotGroupBySample, height = 4, width = 5)
    # check the sample distribution of DimPlot with split.by Sample group by harmony cluster
    AllUrotheliumDimPlotGroupBySample<- DimPlot(obj, group.by = "harmony_clusters", split.by = "Sample")
    # save the plot
    ggsave(paste0(name,"_DimPlotSplitBySample.pdf"),plot = AllUrotheliumDimPlotGroupBySample, height = 4, width = 15)

    # check the number of cells in each sample
    SampleNumber<- table(obj@meta.data$Sample)
    write.csv(SampleNumber,paste0(name,"_Urothelium_SampleNumber.csv"))

    GroupNumber<- table(obj@meta.data$Group)
    write.csv(GroupNumber,paste0(name,"_Urothelium_GroupNumber.csv"))

}


# Some basic information about the object and the metadata

AllUrothelium
AllUrothelium@meta.data %>% head()

# set up the order of Samples levels

AllUrothelium@meta.data$Sample<- ordered(factor(AllUrothelium@meta.data$Sample),levels=c("Vehicle", "Yoda", "LOF", "GOF", "Health","UUO_Day2","UUO_Day4","UUO_Day6","UUO_Day10","UUO_Day14", "GSM3827175_NMU_O_P", "GSM3827176_NMU_O_D"))

# check the sample distribution of DimPlot
AllUrotheliumDimPlotGroupBySample<- DimPlot(AllUrothelium, group.by = "Sample")
# save the plot 
ggsave("AllUrotheliumDimPlotGroupBySample_RunHarmony.pdf",plot = AllUrotheliumDimPlotGroupBySample, height = 4, width = 6)
# check the sample distribution of DimPlot with split.by Sample group by harmony cluster
AllUrotheliumDimPlotSplitBySample<- DimPlot(AllUrothelium, group.by = "seurat_clusters", split.by = "Sample")
# save the plot
ggsave("AllUrotheliumDimPlotSplitBySample_RunHarmony.pdf",plot = AllUrotheliumDimPlotSplitBySample, height = 4, width = 20)

# check the sample distribution of DimPlot with split.by Sample group by harmony cluster
AllUrotheliumDimPlotSplitByGroup<- DimPlot(AllUrothelium, group.by = "seurat_clusters", split.by = "Group")
# save the plot
ggsave("AllUrotheliumDimPlotSplitByGroup_RunHarmony.pdf",plot = AllUrotheliumDimPlotSplitByGroup, height = 4, width = 12)


# check the sample distribution of DimPlot with split.by Sample group by harmony cluster
AllUrotheliumDimPlotGroupByGroup<- DimPlot(AllUrothelium, group.by = "Group")
# save the plot
ggsave("AllUrotheliumDimPlotGroupByGroup_RunHarmony.pdf",plot = AllUrotheliumDimPlotGroupByGroup, height = 4, width = 6)

# check the number of cells in each sample
SampleNumber<- table(AllUrothelium@meta.data$Sample)
write.csv(SampleNumber,"Urothelium_SampleNumber.csv")

GroupNumber<- table(AllUrothelium@meta.data$Group)
write.csv(GroupNumber,"Urothelium_GroupNumber.csv")
```

```{r}
# we used the clustered method to extract the urothelium
table(AllUrothelium@meta.data$Sample)
# we measure the Upk3a, Upk1a, Upk1b and Upk2 positive cell numbers
Upk3aUro <- subset(AllUrothelium, subset = Upk3a > 2 )
Upk1aUro <- subset(AllUrothelium, subset = Upk1a > 2 )
Upk1bUro <- subset(AllUrothelium, subset = Upk1b > 2 )
Upk2Uro <- subset(AllUrothelium, subset = Upk2 > 2 )
table(Upk3aUro@meta.data$sample)
table(Upk1aUro@meta.data$sample)
table(Upk1bUro@meta.data$sample)
table(Upk2Uro@meta.data$sample)
#rm(Upk3aUro, Upk1aUro, Upk1bUro, Upk2Uro)

Krt5Uro <- subset(AllUrothelium, subset = Krt5 > 2 )
Krt14Uro <- subset(AllUrothelium, subset = Krt14 > 2 )

table(Krt5Uro@meta.data$sample)
table(Krt14Uro@meta.data$sample)
# we used the marker method to define the accurate urothelium

Urothelium <-subset(Urothelium_S2, subset = Upk3a > 2 | Upk1a >2 |Upk1b >2 |Upk2 > 2 )

table(Urothelium@meta.data$Group)
# Get the average expression of the genes
UrotheliumAverageExpression<- AverageExpression(object = Urothelium)$RNA
# Get the percentage of the genes
# get all the expressed genes 
UrotheliumExpressedGenes<- rownames(Urothelium)
#UrotheliuGenePercent <- PercentExpressing(seurat_object = Urothelium, features = UrotheliumExpressedGenes)

# other ways
plot <- DotPlot(object = Urothelium, features = UrotheliumExpressedGenes)
plot_data <- plot$data %>% 
  select(avg.exp,pct.exp, id)
# write them into file

write.csv(plot_data, file="Urothelium_ExpressedGenes_All_Ranking.csv")

table(Urothelium@meta.data$sample)
table(UUOProjectObject_F@meta.data$sample)

table(Urothelium_Upk3a_krt5$sample)
dim(Urothelium_Upk3a_krt5)
table(Urothelium@meta.data$sample)
dim(Urothelium)

```

```{r}
# We collected the dataset from the KUDO and Compared the expression of genes from Control
# Correlation of Average Expression
# 1. Compute average expression per cluster (organoid urothelium vs mouse urothelium, and bladder urothelium organoid).
# Pearson or Spearman correlation:
MouseUrothelum <- subset(AllUrothelium, subset = Group == "UUO")

# we only concentrate on the urothelium
MouseUrothelum@meta.data %>% head()
# # 
# MouseUrothelum <- subset(MouseCells, subset = celltype0421 == "Uro")


Idents(MouseUrothelum)
OrganoidUrothelium <- subset(AllUrothelium, subset = Group == "KUDO")
# integrated the layers
DefaultAssay(OrganoidUrothelium) <- "RNA"

OrganoidUrothelium <- JoinLayers(OrganoidUrothelium)



# get the average expression of the genes in the mouse urothelium and organoid urothelium
expr_mat <- GetAssayData(
  MouseUrothelum,
  assay = "RNA",
  layer = "data"
)
mouse_avg <- Matrix::rowMeans(expr_mat)


Oexpr_mat <- GetAssayData(
  OrganoidUrothelium,
  assay = "RNA",
  layer = "data"
)
organoid_avg <- Matrix::rowMeans(Oexpr_mat)

# # Define a threshold for expressed genes
# expression_threshold <- 0  # Adjust this threshold as needed

# # Filter for expressed genes in both datasets

# expressed_genes_mouse <- names(mouse_avg[mouse_avg > expression_threshold])
# expressed_genes_organoid <- names(organoid_avg[organoid_avg > expression_threshold])

# Find shared expressed genes
shared_genes <- intersect(names(mouse_avg), names(organoid_avg))

length(shared_genes)
# Subset the datasets to only include shared expressed genes
mouse_avg_2 <- mouse_avg[shared_genes]
organoid_avg_2 <- organoid_avg[shared_genes]


# calculate the correlation between two datasets
cor_val <- cor(organoid_avg_2, mouse_avg_2, method="spearman")

library(reshape2)
library(ggplot2)

# Combine data into a data frame
data <- data.frame(mouse_avg_2, organoid_avg_2)

# Create the correlation dot plot
MouseUro_OrganoidUro_Cor_Plot<- ggplot(data, aes(x = log10(mouse_avg_2), y = log10(organoid_avg_2))) +
  geom_point(color = "gray", size = 3) +
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  labs(
    title = "Correlation Dot Plot",
    x = "Mouse Avg 2 (log2)",
    y = "Organoid Avg 2 (log2)"
  ) +
  theme_minimal() 
ggsave("MouseUro_OrganoidUro_Cor_Plot.pdf",plot = MouseUro_OrganoidUro_Cor_Plot, height = 4, width = 5)

# we subset the subset(scrna, subset = Group == "Health")

Urothelium_KUDO_MouseH_Obj <- subset(scrna, subset =  Group == "KudoVehicle" |(Group == "Health" & celltype0421 == "Uro"))
Urothelium_KUDO_MouseH_Obj@meta.data %>% head() 
DefaultAssay(Urothelium_KUDO_MouseH_Obj) <- "RNA"
Urothelium_KUDO_MouseH_Obj <- JoinLayers(Urothelium_KUDO_MouseH_Obj)

saveRDS(Urothelium_KUDO_MouseH_Obj, file = "Urothelium_KUDO_MouseH_Obj.rds")



  # +
  # annotate(
  #   "text",
  #   x = min(log10(mouse_avg_2)) + 0.5,
  #   y = max(log10(organoid_avg_2)) - 0.5,
  #   label = paste("Spearman Correlation:", round(cor_val, 2)),
  #   color = "black",
  #   size = 4,
  #   hjust = 0
  # )

# 2. Marker gene expression Overlaping
DotPlot(scrna, features = c("Krt5","Tp63","Upk1a","Upk2","Upk3a")) + RotatedAxis()
# Marker gene overlap (Jaccard index).
# J=∣Markersmouse​∪Markersorganoid​∣∣Markersmouse​∩Markersorganoid​∣​
Idents(Urothelium_KUDO_MouseH_Obj) <- "sample"
levels(Idents(Urothelium_KUDO_MouseH_Obj))
DotPlot(Urothelium_KUDO_MouseH_Obj, features = c("Krt5","Trp63","Upk1a","Upk2","Upk3a")) + RotatedAxis()
```

# method 2: the lineage marker expression correlation
```{r}
# Checking lineage marker expression correlation will tell you whether the urothelial identity program is preserved, independent of stress/proliferation noise.
markers_mouse <- c(
  "Krt5","Krt14","Trp63","Itga6","Krt17",
  "Krt8","Krt18","Foxa1","Gata3","Pparg",
  "Upk1a","Upk1b","Upk2","Upk3b","Upk3a","Krt20"
)
markers_mouse_expanded <- c(
  # Basal
  "Krt5","Krt14","Trp63","Itga6","Krt17","Col17a1","Sox2","Ngfr","Cd44","Lgr5","Pdpn","Epcam",
  # Intermediate
  "Krt8","Krt18","Foxa1","Gata3","Pparg","Grhl3","Upk1a","Upk1b","Upk2","Upk3a","S100a1","Cdh1","Tjp1",
  # Umbrella
  "Krt20","Shh","Plk2","Umod","Lypd3","Cldn8","Cldn4","Ocln"
)
# Subset the datasets to only include shared expressed genes
mouse_avg_markers <- mouse_avg[markers_mouse]
organoid_avg_markers <- organoid_avg[markers_mouse]
cor(mouse_avg_markers,
    organoid_avg_markers,
    method = "pearson")

cor(mouse_avg_markers,
    organoid_avg_markers,
    method = "spearman")

```

```{r}
# we selected the top 1000 variable genes in the mouse urothelium and organoid urothelium
mouse_top1000 <- head(VariableFeatures(MouseUrothelum), 1000)
organoid_top1000 <- head(VariableFeatures(OrganoidUrothelium), 1000)
# calculate the Jaccard index
intersection <- length(intersect(mouse_top1000, organoid_top1000))
union <- length(union(mouse_top1000, organoid_top1000))
jaccard_index <- intersection / union

```


```{r}
# we create a method for the measure the two objects average expresssion correlation
library(Seurat)
library(Matrix)
library(ggplot2)

# Function to compute correlation and generate dot plot
correlation_dotplot <- function(obj1, obj2, assay="RNA", slot="data",
                                genes=NULL, method="spearman",
                                log_scale=TRUE, plot_title=NULL,
                                save_plot=NULL) {
  
  # Get average expression
  expr1 <- Matrix::rowMeans(GetAssayData(obj1, assay=assay, slot=slot))
  expr2 <- Matrix::rowMeans(GetAssayData(obj2, assay=assay, slot=slot))
  
  # If a gene list is provided, subset
  if(!is.null(genes)) {
    shared_genes <- intersect(names(expr1), names(expr2))
    shared_genes <- intersect(shared_genes, genes)
  } else {
    shared_genes <- intersect(names(expr1), names(expr2))
  }
  
  # Subset to shared genes
  expr1_sub <- expr1[shared_genes]
  expr2_sub <- expr2[shared_genes]
  
  # Calculate correlation
  cor_val <- cor(expr1_sub, expr2_sub, method=method)
  
  # Prepare data frame for plotting
  plot_data <- data.frame(expr1_sub, expr2_sub)
  colnames(plot_data) <- c("Dataset1", "Dataset2")
  
  # Apply log10 if requested
  if(log_scale) {
    plot_data$Dataset1 <- log10(plot_data$Dataset1)
    plot_data$Dataset2 <- log10(plot_data$Dataset2)
    xlab <- paste0(deparse(substitute(obj1)), " Avg (log10)")
    ylab <- paste0(deparse(substitute(obj2)), " Avg (log10)")
  } else {
    xlab <- paste0(deparse(substitute(obj1)), " Avg")
    ylab <- paste0(deparse(substitute(obj2)), " Avg")
  }
  
  # Generate plot
  p <- ggplot(plot_data, aes(x=Dataset1, y=Dataset2)) +
    geom_point(color="gray", size=3) +
    geom_smooth(method="lm", color="black", se=FALSE) +
    labs(title = plot_title %||% "Correlation Dot Plot", x=xlab, y=ylab) +
    theme_minimal()
  
  # Save plot if filename provided
  if(!is.null(save_plot)) {
    ggsave(save_plot, plot=p, height=4, width=5)
  }
  
  # Return results
  return(list(correlation=cor_val, plot=p, shared_genes=shared_genes))
}

# Example usage:
res <- correlation_dotplot(MouseUrothelum, OrganoidUrothelium,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_Cor_Plot.pdf")

# Print correlation value
res$correlation

# Display plot
res$plot
length(res$shared_genes)
```

# Define the function II: top 1000 variable genes for jaccard_index 
```{r}
library(Seurat)
library(VennDiagram)

# Function to compute Jaccard index and draw Venn diagram
topVariableGenes_Jaccard <- function(obj1, obj2, assay="RNA", slot="data",
                                     n=1000, plot_title=NULL, 
                                     save_plot=NULL, verbose=TRUE) {
  
  # 1️⃣ Find top variable genes
  hvgs1 <- head(VariableFeatures(FindVariableFeatures(obj1, assay=assay, selection.method="vst", nfeatures=n)), n)
  hvgs2 <- head(VariableFeatures(FindVariableFeatures(obj2, assay=assay, selection.method="vst", nfeatures=n)), n)
  
  if(verbose) {
    message("Number of HVGs selected: ", length(hvgs1), " and ", length(hvgs2))
  }
  
  # 2️⃣ Compute Jaccard index
  shared <- intersect(hvgs1, hvgs2)
  union_genes <- union(hvgs1, hvgs2)
  jaccard_index <- length(shared) / length(union_genes)
  
  if(verbose) {
    message("Shared genes: ", length(shared))
    message("Union genes: ", length(union_genes))
    message("Jaccard index: ", round(jaccard_index, 3))
  }
  
  # 3️⃣ Generate Venn diagram
  venn_plot <- venn.diagram(
    x = list(Object1 = hvgs1, Object2 = hvgs2),
    filename = NULL,
    fill = c("skyblue", "pink"),
    alpha = 0.5,
    cex = 1.5,
    cat.cex = 1.2,
    main = plot_title %||% paste0("Top ", n, " HVGs Venn Diagram")
  )
  
  # Save plot if needed
  if(!is.null(save_plot)) {
    png(save_plot, width = 600, height = 600)
    grid.draw(venn_plot)
    dev.off()
  }
  
  # Return results
  return(list(hvgs_object1=hvgs1,
              hvgs_object2=hvgs2,
              shared_genes=shared,
              jaccard_index=jaccard_index,
              venn_plot=venn_plot))
}

# Example usage
Urotheliumres <- topVariableGenes_Jaccard(MouseUrothelum, OrganoidUrothelium,
                                n=2000,
                                plot_title="Mouse vs Organoid Top HVGs",
                                save_plot="Mouse_Organoid_HVGs_Venn.png")

# View Jaccard index
Urotheliumres$jaccard_index
Urotheliumres$venn_plot
```

# #####################
# UPK lineage
# #####################
```{r}
# for the objects, we only selected the cells that showe expression of Upk3a, Upk1a, Upk1b and Upk2.
UPKUrotheliumMouse<- subset(MouseUrothelum, Upk3a > 1 | Upk1a > 1 | Upk1b > 1 | Upk2 > 1)
# we also selected the cells that showe expression of Upk3a, Upk1a, Upk1b and Upk2 in the organoid urothelium
UPKUrotheliumOrganoid<- subset(OrganoidUrothelium, Upk3a > 1 | Upk1a > 1 | Upk1b > 1 | Upk2 > 1)

# print the correlation:
UPKres <- correlation_dotplot(UPKUrotheliumMouse, UPKUrotheliumOrganoid,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid UPK lineage Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_UPKPos_Cor_Plot.pdf")

# Print correlation value
UPKres$correlation

# Display plot
UPKres$plot

length(UPKres$shared_genes)

# # Venn diagram and Jaccard index for the top variable genes in the UPK positive cells
UPKtopvariable <- topVariableGenes_Jaccard(UPKUrotheliumMouse, UPKUrotheliumOrganoid,
                                n=2000,
                                plot_title="Mouse vs Organoid Top UPKs",
                                save_plot="Mouse_Organoid_top2000_UPKPos_Venn.png")
                                
UPKtopvariable$venn_plot 
UPKtopvariable$jaccard_index                             
```

# #####################
# KRT5, Krt14 lineage
# #####################
```{r}
# for the objects, we only selected the cells that showe expression of Krt5 and Krt14.
KRT5UrotheliumMouse<- subset(MouseUrothelum, Krt5 > 1 |Krt14 > 1)
# we also selected the cells that showe expression of Krt5 and Krt14.
KRT5UrotheliumOrganoid<- subset(OrganoidUrothelium, Krt5 > 1 |Krt14 > 1))

# print the correlation:
KRTres <- correlation_dotplot(KRT5UrotheliumMouse, KRT5UrotheliumOrganoid,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid KRT5 lineage Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_KRT5Pos_Cor_Plot.pdf")

# Print correlation value
KRTres$correlation
KRTres$plot
length(KRTres$shared_genes)
# Venn diagram and Jaccard index for the top variable genes in the KRT5 and KRT14 positive cells
KRTtopvariable <- topVariableGenes_Jaccard(KRT5UrotheliumMouse, KRT5UrotheliumOrganoid,
                                n=2000,
                                plot_title="Mouse vs Organoid Top KRT",
                                save_plot="Mouse_Organoid_top2000_KRTPos_Venn.png")
KRTtopvariable$jaccard_index   
```

# ##############
# UPK,KRT lineage
# ##############
```{r}
# we selected the KRT and UPK positive cells in the mouse urotheloium

UPKKRTUrotheliumMouse<- subset(MouseUrothelum, subset = (Upk3a > 1 | Upk1a > 1 | Upk1b > 1| Upk2 > 1 | Krt5 > 1 |Krt14 > 1)) 

# we also selected the KRT and UPK positive cells in the organoid urotheloium
UPKKRTUrotheliumOrganoid<- subset(OrganoidUrothelium,(Upk3a > 1 | Upk1a > 1 | Upk1b > 1| Upk2 > 1 | (Krt5 > 1 |Krt14 > 1)) 
# print the correlation:
UPKKRTres <- correlation_dotplot(UPKKRTUrotheliumMouse, UPKKRTUrotheliumOrganoid,
                           assay="RNA", slot="data",
                           method="spearman",
                           log_scale=TRUE,
                           plot_title="Mouse vs Organoid UPK+ or KRT+ Urothelial Correlation",
                           save_plot="MouseUro_OrganoidUro_UPKKRT5Pos_Cor_Plot.pdf")

# Print correlation value
UPKKRTres$correlation
UPKKRTres$plot
length(UPKKRTres$shared_genes)

# Venn diagram and Jaccard index for the top variable genes in the UPK+ or KRT+ cells
UPKKRTtopvariable <- topVariableGenes_Jaccard(UPKKRTUrotheliumMouse, UPKKRTUrotheliumOrganoid,
                                n=2000,
                                plot_title="Mouse vs Organoid Top UPK KRT",
                                save_plot="Mouse_Organoid_top2000_UPKKRTPos_Venn.png")
```


```{r}
# check the annotation of scrna
Idents(scrna)
scrna@meta.data %>% head()
scrna@meta.data$celltype0421
```

# ############### ##############
# ############### ##############
# Reclustering the Urothelium_KUDO_MouseH_Obj to check the overlapping of the two datasets
# ############### ##############
# ############### ##############

```{r}
Urothelium_KUDO_MouseH_Obj<- readRDS("Urothelium_KUDO_MouseH_Obj.rds")
# we recluster the Urothelium_KUDO_MouseH_Obj
Urothelium_KUDO_MouseH_Obj <- Urothelium_KUDO_MouseH_Obj %>%  
  NormalizeData() %>% 
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
  ScaleData() %>% 
  SCTransform(vars.to.regress = c("percent.mt"))

## find most vaible features across sample to integrate


Urothelium_KUDO_MouseH_Obj <-  RunPCA(scrna, features = VariableFeatures(object = Urothelium_KUDO_MouseH_Obj),assay = "SCT",npcs=20)
#scrna <-readRDS(file = "KudoUUOComparision_Scrna_PCA.rds")
# then use the Harmony to remove batches
Urothelium_KUDO_MouseH_Obj <- RunHarmony(Urothelium_KUDO_MouseH_Obj, group.by.vars = 'sample', dims.use = 1:20, plot_convergence= FALSE, reduction = "pca", assay.use = "SCT", reduction.save = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_Harmony.rds")
Urothelium_KUDO_MouseH_Obj <- RunUMAP(Urothelium_KUDO_MouseH_Obj, dims = 1:20,reduction = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_RunUMAP.rds")
Urothelium_KUDO_MouseH_Obj <- FindNeighbors(Urothelium_KUDO_MouseH_Obj, dims = 1:20, reduction = "harmony")
#saveRDS(scrna, file = "KudoUUOComparision_Scrna_FindNeighbor.rds")
Urothelium_KUDO_MouseH_Obj <- FindClusters(Urothelium_KUDO_MouseH_Obj, resolution = 0.2)

OverlappingDimPlot<- DimPlot(Urothelium_KUDO_MouseH_Obj)
ggsave("OverlappingDimPlot.pdf",plot = OverlappingDimPlot, height = 4, width = 5)
OverlappingDimPlotSplit<- DimPlot(Urothelium_KUDO_MouseH_Obj, split.by = "sample")
ggsave("OverlappingDimPlotSplit.pdf",plot = OverlappingDimPlotSplit, height = 4, width = 9)

saveRDS(Urothelium_KUDO_MouseH_Obj, file = "Urothelium_KUDO_MouseH_Obj_Reclustered.rds")

Urothelium_KUDO_MouseH_Obj <- readRDS("Urothelium_KUDO_MouseH_Obj_Reclustered.rds")
markers_mouse_expanded <- c(
  # Basal
  "Krt5","Krt14","Trp63","Itga6","Krt17","Col17a1","Sox2","Ngfr","Cd44","Lgr5","Pdpn","Epcam",
  # Intermediate
  "Krt8","Krt18","Foxa1","Gata3","Pparg","Grhl3","Upk1a","Upk1b","Upk2","Upk3a","S100a1","Cdh1","Tjp1",
  # Umbrella
  "Krt20","Shh","Plk2","Umod","Lypd3","Cldn8","Cldn4","Ocln"
)

Idents(Urothelium_KUDO_MouseH_Obj) <- "sample"
UrothelialMarkersDotPlot<- DotPlot(Urothelium_KUDO_MouseH_Obj, features = markers_mouse_expanded) + RotatedAxis()
ggsave("UrothelialMarkersDotPlot_Urothelim_KUDO_MouseH_Reclustered.pdf", plot=UrothelialMarkersDotPlot, height=3, width=12)
markers_mouse_expanded_withothers<- c("Krt5","Krt14","Trp63",
  "Krt8","Krt18","Foxa1",
  "Upk1a","Upk1b","Upk2","Upk3b","Upk3a","Krt20",
  "Lrp2","Slc34a1","Slc13a3", # Proximal tubule cells (PT)
  "Slc12a3","Pvalb","Wnk1",# Distal connecting tublue (DCT)
"Umod","Slc12a1","Cldn10", # Thick ascending limb cells (TAL)
                      "Klk1","Slc8a1","Calb1", # Connecting tubule (CNT)
                     "Slc14a2","Fst","Bst1", # descending thin  limb cells (DTL)
                     #"Akr1b1","Sh3gl3","Prox1", # ascending thin limb cells (ATL)
                     #"Enox1","Thsd4", #Macula densa
                      "Aqp2","Hsd11b2","Scnn1g", # principal cells (PC)
                    "Atp6v1g3","Aqp6","Slc26a7", # Intercalated cells
                    "Cdh5","Igfbp3","Pecam1", # Endothelium
                    "Col3a1","Fbn1","Lum","Col1a1", # Fibroblast
                    "Ireb2","Alas2", # Plasmacytoid dentritic cells
                    "S100a8","S100a9","Il1b", # Neutrophil
                    "C1qa","C1qb","Aif1", # Macrophage
                    "Cxcr6","Cd247","Cd3e", # T cells
                    "Igkc","Cd79a","Cd79b", # B cells
                    "Ccl5","Nkg7","Cd7" # NK/Cd8 cells
  
  )
KidneyMarkersDotPlot<- DotPlot(Urothelium_KUDO_MouseH_Obj, features = markers_mouse_expanded_withothers) + RotatedAxis()
ggsave("KidneyMarkersDotPlot_Urothelim_KUDO_MouseH_Reclustered.pdf", plot=KidneyMarkersDotPlot, height=4, width=20)

# check the average expression of the markers in the mouse urothelium and organoid urothelium

avg_exp <- AverageExpression(
  object = Urothelium_KUDO_MouseH_Obj,
  features = markers_mouse_expanded_withothers,
  group.by = "sample"   # or "seurat_clusters", "condition"
)
avg_exp%>% head()
# set up the order of the columns in the same order as the markers
aveg_exp_df <- as.data.frame(avg_exp$RNA)


# identify missing genes
missing_genes <- setdiff(markers_mouse_expanded_withothers, rownames(aveg_exp_df))
if (length(missing_genes) > 0) {
  zero_mat <- matrix(
    0,
    nrow = length(missing_genes),
    ncol = ncol(aveg_exp_df),
    dimnames = list(missing_genes, colnames(aveg_exp_df))
  )
  
  aveg_exp_df <- rbind(aveg_exp_df, zero_mat)
}

aveg_exp_df$Gene <- rownames(aveg_exp_df)

# change to the long type
aveg_exp_df_long <- aveg_exp_df %>%
  pivot_longer(cols = -Gene, names_to = "Sample", values_to = "AverageExpression")

library(ggplot2)
# set the order of the genes in the same order as the markers
aveg_exp_df_long$Gene <- factor(aveg_exp_df_long$Gene, levels = markers_mouse_expanded_withothers)

AvereageExpressionBarPlot <-
ggplot(aveg_exp_df_long, aes(x = Gene, y = AverageExpression, fill= Sample)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  theme_classic(base_size = 14) +
  labs(y = "Average Expression", x = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )
ggsave("AvereageExpressionBarPlot_Urothelim_KUDO_MouseH_Reclustered.pdf", plot=AvereageExpressionBarPlot, height=4, width=15)

```